Spoken Corpus of Karel Mako\v{n}

### 2018-04-13 2018-clarin-pisa/makon-data.tex
more clarin text

\usepackage{covington} % if needed, for linguistic examples

@@

\begin{document}
\maketitle
\begin{abstract}
  We present a detailed view on the spoken corpus of Karel Mako\v{n}. This paper
  briefly outlines the system for processing the corpus, which has been
  published in a previous work, and elaborates on the corpus. A baseline for
  topical analysis is presented.
\end{abstract}

@@

His steep spiritual path began in early childhood when he was subject to
repeated arm surgery without anesthesy. The pain the child suffered made him
escape mentally from inside his body, which led him to what he calls ``the
yellow light''. As a result, he felt obliged always to {\em do the right thing}.

This culminated in the age of 17, in a perceived life crissis, when he read in a
book that ``this life is the bridge to the eternal life''. The thought
captivated him and gave him a new life direction, which led to several
ecstasies.

The next culmination happened in 1939, after nine years of constant prayer for a
stronger love to God, when he was deported to the nazi concentration camp in
Sachsenhausen, as a university student. Here, he was promised certain death by a
Gestapo warden. As the nazi was approaching K.M. to kill him, he gave up his
life completely to God and the attacker unexplainably turned around in awe and
fled.

@@

The mystical, religious and spiritual nature of Mako\v{n}'s talks put K.M. in
direct antagonism with the communist regime. The opposition was strictly
unidirectional though, as Mako\v{n} also understood the purpose of the regime
and was far from wasting his energy on a fight against it. His way of dealing
with the conflict was by staying out of wide popularity. Hence, most of the
talks have been given during regular meetings in various places of
Czechoslovakia and later the Czech Republic, in a close circle of friends.

One of the fortunate consequences of the seclusion of the talks is that there is
very little background noise on most of the recordings. Also, most of them
% TODO: kolik je od Elgra?
have been taken by a single person who used the available state-of-the-art
amateur recording technology and methods. He was also very systematic in
labeling and archiving the recordings. As a result, the majority of the base for
digitization was a set of reel-to-reel tapes labeled with ordinal indexes and a
set of cassettes labeled by year and ordinal index. The reel-to-reel tapes have
been acquired in a generous speed of 9cm/second, which granted them a very good
shield against the inevitable aging.

@@

The speech recordings are available through two sources. First and foremost,
it is in the Clarin repository, which guarantees it will be easily available no
matter what fate this project will take.

Secondly, I have created a web application for listening to the recordings
directly in the browser. The web application is also capable of displaying the
transcription synchronously. When there is an error in the transcription (which
is frequently so in case of automatically acquired ones), the user can submit a
correction.

@@

We only have a single speaker, a restricted domain, relatively consistent
conditions, an extensive collection of the author's original books,
translations, comments and letters to build a fitting language model. So
even very little training data should give satisfactory results and the error
rate should drop rapidly as the training data would increase.

Well, the error rate did drop as the training data increased but the accuracy of
the speech recognition was more of a challenge than I had thought. For the most
part, the users were actually using the web interface to provide a new
transcription from scratch, ignoring the way too erroneous output of the ASR,
instead of merely correcting single errors, as was meant.

Still, the community managed to provide over 66 hours of high-quality manual
transcriptions, which have been used gradually as training data. Now, the
acoustically intact recordings are automatically transcribed well enough to
allow reading along and to do full-text search.

@@

The varying acoustic quality is the most serious obstacle to the quality of the
automatic transcription. But it is also detrimental to the manual transcription.

To mitigate this, the web interface features since lately a user-controllable
graphical equalizer, which can help especially in the case of narrow-band noise.

@@

Of course, there are exceptions, like saying ``the capital of Japan'' instead of
``Tokyo'' but firstly, I chose terms where this is not likely to happen and
secondly, this can only lower the recall of our baseline score, not the
precision, and we have no way of telling the recall given no annotated data
for this task.\footnote{The 50-minute test set is way too small for this kind of
task. Another source could be manual notes available to some of the recordings
but preparing them for this purpose requires some more effort.}

I have chosen the following example topics:
1. \emph{Lazarus},
2. \emph{Mithra, Mithraism},
3. \emph{Satan},
4. \emph{St. Teresa},
5. \emph{fairy-tale}.

I chose fairy-tale (Czech: \emph{poh\'{a}dka}) as a recurring topic with a
unifying word, for an example of a non-named-entity.

The search for stems of these words in the automatic part of the transcription
yielded results summed up in Table~\ref{tab:topicsearch}. For terms where there
were more than 20 hits, I have checked the first 20 but each from a different
file.

\begin{table}[htpb]
\begin{center}
\begin{tabular}{|l|l|r|r|r|}
\hline
Term & Query & hits in automatic transcription & hits in manual transcription & Precision \\
\hline
Lazarus & \texttt{lazar.*} & 16 & 14 & 8/10   \\
Mithra & \texttt{mithra.*|mitra.*} & 0 & 0 & n/a   \\
Satan & \texttt{satan.*} & 308 & 129 & 16/20   \\
St. Teresa & \texttt{terez.*} & 882 & 91 & 15/20   \\
fairy-tale & \texttt{pohádk.*} & 216 & 55 & 18/20   \\
\hline
\multicolumn{4}{|l|}{average precision} & 81.25\%\\
\hline
\end{tabular}
\caption{Text search hits in the automatically-acquired part of the
transcription}\label{tab:topicsearch}
\end{center}
\end{table}

@@

>>>>

\usepackage{covington} % if needed, for linguistic examples
\usepackage{makecell}

@@

\begin{document}
\maketitle
\begin{abstract}
  We present the spoken corpus of Karel Mako\v{n}. This paper
  briefly outlines the system for processing the corpus, which has been
  published in a previous work, and elaborates on the corpus. A baseline for
  topical analysis is presented.
\end{abstract}

@@

His steep spiritual path began in early childhood when he was subject to
repeated arm surgery without anesthesy. The pain the child suffered made him
escape mentally from inside his body, which led him to what he calls ``the
yellow light''.

His path culminated in 1939, after nine years of constant prayer for a
stronger love to God, when he was deported to the nazi concentration camp in
Sachsenhausen, as a university student. Here, he was promised certain death by a
Gestapo warden. As the nazi was approaching K.M. to kill him, he gave up his
life completely to God and the attacker unexplainably turned around in awe and
fled.

@@

The mystical, religious and spiritual nature of Mako\v{n}'s talks put K.M. in
direct antagonism with the communist regime. The opposition was strictly
unidirectional though, as Mako\v{n}
was far from wasting his energy on a fight against it. His way of dealing
with the conflict was by staying out of wide popularity. Hence, most of the
talks have been given during regular meetings in various places of
Czechoslovakia and later the Czech Republic, in a close circle of friends.

One of the fortunate consequences of the seclusion of the talks is that there is
very little background noise on most of the recordings. Also, most of them
% TODO: kolik je od Elgra?
have been taken by a single person who used the available state-of-the-art
amateur recording technology and methods. This person was also very systematic
in labeling and archiving the recordings. As a result, the majority of the base
for digitization was a set of reel-to-reel tapes labeled with ordinal indexes
and a set of cassettes labeled by year and ordinal index. The reel-to-reel tapes
have been recorded in a generous speed of 9cm/second, which granted them a very
good shield against the inevitable aging.

@@

The speech recordings are available through two sources. Firstly,
it is in the Clarin repository, which guarantees it will be easily available no
matter what fate this project will take.

Secondly, I have created a web application for listening to the recordings
directly in the browser. The web application is also capable of displaying the
transcription synchronously. When there is an error in the transcription (and
the automatic transcription has plenty of errors), the user can submit a
correction.

@@

The community managed to provide over 66 hours of high-quality manual
transcriptions, which have been used gradually as training data. Now, the
acoustically intact recordings are automatically transcribed well enough to
allow reading along and to do full-text search.

@@

The varying acoustic quality is the most serious obstacle to the quality of the
automatic transcription. But it is also detrimental to the manual transcription.
To mitigate this, the web interface features since lately a user-controllable
graphical equalizer, which can help especially in the case of narrow-band noise.

@@

Of course, there are exceptions, like saying ``the capital of Japan'' instead of
``Tokyo'' but firstly, I chose terms where this is not likely to happen and
secondly, this can only lower the recall of our baseline score, not the
precision, and we have no way of telling the recall given no annotated data
for this task anyway.\footnote{The 50-minute test set is way too small for this
kind of task. Another source could be manual notes available to some of the
recordings but preparing them for this purpose requires some more effort.}

I have chosen the following example topics:
1. \textbf{Lazarus},
2. \textbf{Mithra, Mithraism},
3. \textbf{Satan},
4. \textbf{St. Teresa},
5. \textbf{fairy-tale}.
I chose fairy-tale (Czech: \emph{poh\'{a}dka}) as a recurring topic with a
unifying word, for an example of a non-named-entity.

The search for stems of these words in the transcription
yielded results summed up in Table~\ref{tab:topicsearch}. For terms where there
were more than 20 hits, I have checked the first 20 but each from a different
file.

\begin{table}[htpb]
\begin{center}
\begin{tabular}{|l|l|r|r|r|r|}
\hline
Term &
Query &
\makecell{hits in automatic\\
transcription} &
\makecell{expected hits\\
in automatic\\
transcription} &
\makecell{hits in manual\\
transcription} &
Precision \\
\hline
Lazarus & \texttt{lazar.*} & 16 & 171 & 14 & 8/10   \\
Mithra & \texttt{mithra.*|mitra.*} & 0 & 0 & 0 & n/a   \\
Satan & \texttt{satan.*} & 308 & 1575 & 129 & 16/20   \\
St. Teresa & \texttt{terez.*} & 882 & 1111 & 91 & 15/20   \\
fairy-tale & \texttt{pohádk.*} & 216 & 671 & 55 & 18/20   \\
\hline
\multicolumn{5}{|l|}{average precision} & 81.25\%\\
\hline
\end{tabular}
\caption{Text search hits in the transcription; precision only calculated from
automatic transcription}\label{tab:topicsearch}
\end{center}
\end{table}

@@

7.6\% of the total words in the corpus are manually transcribed.
